---
title: "Ex_6_Yuval_Ben_Eliahu"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(pheatmap)
library(cluster)
library(Rtsne)

expFile=system.file("extdata",
                    "leukemiaExpressionSubset.rds",
                    package="compGenomRData")

mat=readRDS(expFile)

scaledMat <- scale(mat)

logMat <- log2(mat)

scaledLogMat <- scale(logMat)

annotation_col = data.frame(
                    LeukemiaType =substr(colnames(mat),1,3))
rownames(annotation_col)=colnames(mat)
```

For the transformed and untransformed data sets used in the exercise above, use the silhouette for deciding number of clusters using hierarchical clustering. [Difficulty: Intermediate/Advanced]
```{r}
Ks=sapply(2:7,
function(i)
summary(silhouette(pam(t(mat),k=i)))$avg.width)
plot(2:7,Ks,xlab="k",ylab="av. silhouette",type="b",
pch=19)
hcl_mat=hclust(dist(t(mat)))
clu_mat=cutree(hcl_mat,k=4)
table(clu_mat)

Ks=sapply(2:7,
function(i)
summary(silhouette(pam(t(scaledMat),k=i)))$avg.width)
plot(2:7,Ks,xlab="k",ylab="av. silhouette",type="b",
pch=19)
hcl_scaledMat=hclust(dist(t(scaledMat)))
clu_scaledMat=cutree(hcl_scaledMat,k=4)
table(clu_scaledMat)


Ks=sapply(2:7,
function(i)
summary(silhouette(pam(t(logMat),k=i)))$avg.width)
plot(2:7,Ks,xlab="k",ylab="av. silhouette",type="b",
pch=19)
hcl_logMat=hclust(dist(t(logMat)))
clu_logMat=cutree(hcl_logMat,k=3)
table(clu_logMat)


Ks=sapply(2:7,
function(i)
summary(silhouette(pam(t(scaledLogMat),k=i)))$avg.width)
plot(2:7,Ks,xlab="k",ylab="av. silhouette",type="b",
pch=19)
hcl_scaledLogMat=hclust(dist(t(scaledLogMat)))
clu_scaledLogMat=cutree(hcl_scaledMat,k=4)
table(clu_scaledLogMat)

```

In this exercise we use the Rtsne() function on the leukemia expression data set. Try to increase and decrease perplexity t-sne, and describe the observed changes in 2D plots. [Difficulty: Beginner]
```{r}
set.seed(30) 

tsne_out5 <- Rtsne(t(mat),perplexity = 5) 
plot(tsne_out5$Y,col=as.factor(annotation_col$LeukemiaType),pch=19)

tsne_out10 <- Rtsne(t(mat),perplexity = 10) 
plot(tsne_out10$Y,col=as.factor(annotation_col$LeukemiaType),pch=19)

tsne_out19 <- Rtsne(t(mat),perplexity = 19) 
plot(tsne_out19$Y,col=as.factor(annotation_col$LeukemiaType),pch=19)

tsne_out1 <- Rtsne(t(mat),perplexity = 1) 
plot(tsne_out1$Y,col=as.factor(annotation_col$LeukemiaType),pch=19)

legend("bottomleft",
legend=unique(annotation_col$LeukemiaType),
fill =palette("default"),
border=NA,box.col=NA)

#With low perplexity of 1, the points form many close small clusters with high local variation. 
#with high perplexity of 19 the points begin to merge into a few big clusters with high global variation.
#with medium perplexity between 5 and 10 the clusters are mostly well defined and meaningful.
```
